---
title: "Learning Interactive Plotting Packages"
output: 
  html_document: 
    theme: spacelab
    toc: yes
---
```{r setup, include=FALSE}

#packages ----
library(dplyr)
library(ggplot2)
library(shiny)
library(highcharter)
library(dygraphs)
library(plotly)
library(flexdashboard)
library(lubridate)
library(tidyr)
library(reshape2)
library(RCurl)
library(rvest)
library(scales)
library(astsa)
library(ggthemes)
library(reshape)
library(car)
library(forecast)
library(fGarch)
library(xts)

#loads and formats shooting data----

#shooting data
data <- read.csv("https://raw.githubusercontent.com/washingtonpost/data-police-shootings/master/fatal-police-shootings-data.csv")
data$date <- ymd(data$date)
data$age <- as.numeric(data$age)
data$state.flag <- "other"
data$state.flag[data$state == "PA"] <- "PA"
data$state.flag <- as.factor(data$state.flag)
data$city.flag <- "other"
data$city.flag[data$city== "Philadelphia"] <- "PHL"
data$city.flag <- as.factor(data$city.flag)
data$month <- month(data$date, label = TRUE)
data$year <- year(data$date)
data$year <- as.factor(data$year)
data$day <- wday(data$date, label = TRUE, abbr = TRUE)
data$count <- 1
data$age[is.na(data$age)] <- 0

#re-codes race as white, black hispanic, other
data$race2 <- "Other"
data$race2[data$race == "W"] <- "White"
data$race2[data$race == "B"] <- "Black"
data$race2[data$race == "H"] <- "Hispanic"
data$race2[data$race == " "] <- "Unknown"
data$race2 <- as.factor(data$race2)

#creates age buckets
data$age2 <- "Under 18"
data$age2[18 <= data$age & data$age < 30] <- "18 to 29"
data$age2[30 <= data$age & data$age < 45] <- "30 to 44"
data$age2[45 <= data$age] <- "45 and Older"
data$age2[data$age == 0] <- "Unknown"
data$age2 <- as.factor(data$age2)

#loads and checks structure
#View(data)
#str(data)

#state population data
states <- read.csv("http://www.census.gov/popest/data/state/totals/2015/tables/NST-EST2015-01.csv", header = FALSE)
states <- states[-c(1, 2, 3, 4, 5, 6, 7, 8, 9, 61, 62, 63, 64, 65, 66, 67), -c(2,3,4,5,6,7,8)] #removes unwanted rows and columns, keeps only 2015 population and state name
names(states)[1] <- "state"
names(states)[2] <- "pop"
states$state <- sub('.', '', states$state) #removes "." at beginning of state names
states$pop <- gsub(',', '', states$pop) #removes commas from population column
states$state <- state.abb[match(states$state,state.name)] #adds state abbreviations
states[9,1] <- "DC" #Adds Washington DC abbreviation
states$state <- as.factor(states$state)
states$pop <- as.numeric(states$pop)

#loads and checks structure
#View(states)
#str(states)

#joins shooting and population data
data <- left_join(data, states, by = "state")
data$state <- as.factor(data$state)
#str(data)

#creates dataframe of list of dates to fill in date gaps
date.ref <- data.frame(date=seq(as.Date("2015-01-01"), as.Date("2016-01-15"), by="days"))
date.ref$date <- ymd(date.ref$date)

#joins data and date.ref dataframes and replaces na with 0's
data <- merge(data,date.ref,by.x='date',by.y='date',all.x=T,all.y=T)
data$count[is.na(data$count)] <- 0

#loads in and formats passenger, employment and earnings data ----
x.2007 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2007.csv")
data.2007 <- read.csv(text = x.2007)
x.2008 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2008.csv")
data.2008 <- read.csv(text = x.2008)
x.2009 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2009.csv")
data.2009 <- read.csv(text = x.2009)
x.2010 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2010.csv")
data.2010 <- read.csv(text = x.2010)
x.2011 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2011.csv")
data.2011 <- read.csv(text = x.2011)
x.2012 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2012.csv")
data.2012 <- read.csv(text = x.2012)
x.2013 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2013.csv")
data.2013 <- read.csv(text = x.2013)
x.2014 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2014.csv")
data.2014 <- read.csv(text = x.2014)
x.2015 <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/2015.csv")
data.2015 <- read.csv(text = x.2015)

#data.2007 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2007.csv")
#data.2008 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2008.csv")
#data.2009 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2009.csv")
#data.2010 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2010.csv")
#data.2011 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2011.csv")
#data.2012 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2012.csv")
#data.2013 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2013.csv")
#data.2014 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2014.csv")
#data.2015 <- read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/2015.csv")

data.2007[5] <- NULL
data.2007$year <- "2007"
data.2008[5] <- NULL
data.2008$year <- "2008"
data.2009[5] <- NULL
data.2009$year <- "2009"
data.2010[5] <- NULL
data.2010$year <- "2010"
data.2011[5] <- NULL
data.2011$year <- "2011"
data.2012[5] <- NULL
data.2012$year <- "2012"
data.2013[5] <- NULL
data.2013$year <- "2013"
data.2014[5] <- NULL
data.2014$year <- "2014"
data.2015[5] <- NULL
data.2015$year <- "2015"

pax <- rbind(data.2007, data.2008, data.2009, data.2010, data.2011, data.2012, data.2013, data.2014, data.2015)
names(pax) <- tolower(names(pax))
phl <- pax %>% dplyr::filter(origin == "PHL") %>% group_by(year, month) %>% summarise("pax" = sum(passengers))
phl$month2 <- month(phl$month, label = TRUE)
phl$month <- NULL
names(phl)[3] <- "month"
phl$date <- paste(phl$month, "1,", phl$year, sep=" ")
phl$date <- mdy(phl$date)
#write.csv(phl, file = "~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/phl.csv")

#emp 
x.emp <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/phl_monthly_emp.csv")
emp <- read.csv(text = x.emp)
#emp <- (read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/phl_monthly_emp.csv"))

emp <-  melt(emp, id.vars = c("Year"))
names(emp)[1] <- "year"
names(emp)[2] <- "month"
names(emp)[3] <- "emp"
emp$year <- as.character(emp$year)
emp$emp <- emp$emp * 1000
emp$month2[emp$month == "January"] <- 1
emp$month2[emp$month == "February"] <- 2
emp$month2[emp$month == "March"] <- 3
emp$month2[emp$month == "April"] <- 4
emp$month2[emp$month == "May"] <- 5
emp$month2[emp$month == "June"] <- 6
emp$month2[emp$month == "July"] <- 7
emp$month2[emp$month == "August"] <- 8
emp$month2[emp$month == "September"] <- 9
emp$month2[emp$month == "October"] <- 10
emp$month2[emp$month == "November"] <- 11
emp$month2[emp$month == "December"] <- 12
emp$month <- month(emp$month2, label = TRUE)
emp$month2 <- NULL
emp <- emp %>% group_by(year, month) %>% summarise("emp" = sum(emp))
emp <- emp[-c(105, 106, 107, 108), ]

#earnings
x.earn <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/phl_avg_wk_earn.csv")
earnings <- read.csv(text = x.earn)
#earnings <- (read.csv("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/phl_avg_wk_earn.csv"))

earnings <-  melt(earnings, id.vars = c("Year"))
names(earnings)[1] <- "year"
names(earnings)[2] <- "month"
names(earnings)[3] <- "earnings"
earnings$year <- as.character(earnings$year)
earnings$earnings <- earnings$earnings * 4
earnings$month2[earnings$month == "January"] <- 1
earnings$month2[earnings$month == "February"] <- 2
earnings$month2[earnings$month == "March"] <- 3
earnings$month2[earnings$month == "April"] <- 4
earnings$month2[earnings$month == "May"] <- 5
earnings$month2[earnings$month == "June"] <- 6
earnings$month2[earnings$month == "July"] <- 7
earnings$month2[earnings$month == "August"] <- 8
earnings$month2[earnings$month == "September"] <- 9
earnings$month2[earnings$month == "October"] <- 10
earnings$month2[earnings$month == "November"] <- 11
earnings$month2[earnings$month == "December"] <- 12
earnings$month <- month(earnings$month2, label = TRUE)
earnings$month2 <- NULL
earnings <- earnings %>% group_by(year, month) %>% summarise("earnings" = sum(earnings))
earnings <- earnings[-c(105, 106, 107, 108), ]

#jet fuel SOURCE: U.S. Gulf Coast Kerosene-Type Jet Fuel Spot Price FOB Dollars per Gallon
x.fuel <- getURL("https://raw.githubusercontent.com/brndngrhm/time_series_and_forecasting/master/project/phl/data/fuel.csv")
fuel <- read.csv(text = x.fuel)
fuel$date <- mdy(fuel$date)
fuel <- fuel %>% dplyr::filter(date > "2006-12-01") %>% dplyr::filter(date < "2015-09-01")
fuel$year <- year(fuel$date)
fuel$month <- month(fuel$date, label=T)
fuel$year <- as.character(fuel$year)
fuel$date <- NULL

#join data and save .rda
socio <- left_join(emp, earnings, by = c("year", "month"))
phl <- left_join(phl,fuel, by = c("year", "month"))
phl <- left_join(phl, socio, by = c("year", "month"))
phl$year <- as.factor(phl$year)
phl <- phl %>% select(date, year, month, pax, price, emp, earnings)




```

-----------------------------------------------------------------------
## Highcharter {.tabset}

### Basic Bar Chart
```{r,echo = T, fig.width=15, fig.height=7}
#plot of shootings by state
state <- data %>% dplyr::filter(year == "2015") %>% group_by(state, pop) %>% summarise(total = n())

(state.plot <- highchart() %>%
  hc_add_series(name="Shootings", data = ((state$total/state$pop)*100), type = "column") %>%
  hc_xAxis(categories = state$state, color ="#2ecc71") %>%
  hc_title(text="Fatal Shootings by Police", align="left") %>%
  hc_subtitle(text="Per Capita, 2015", align="left") %>%
   hc_yAxis(title=" ") %>%
   hc_exporting(enabled = TRUE) %>%
   hc_credits(enabled = TRUE, # add credits
             text = "Washington Post",
             href = "https://github.com/washingtonpost/data-police-shootings",
             align = "left") %>% 
  hc_add_theme(hc_theme_gridlight()) %>%
  hc_legend(enabled = FALSE))
  
```

### Bar Chart with Diff colors

```{r, echo = T}
#plot of shootings by race
race2 <- data %>% dplyr::filter(year == "2015") %>% group_by(race2) %>% summarise(total = n())

(race.plot <- highchart() %>% 
  hc_add_series(name="Shootings", data=race2$total, type = "column", colorByPoint=T) %>%
  hc_xAxis(categories = race2$race2) %>%
  hc_title(text="Fatal Shootings by Police", align="left") %>%
  hc_subtitle(text="Broken out by race, 2015", align="left") %>%
   hc_yAxis(title=" ") %>%
   hc_exporting(enabled = TRUE) %>%
   hc_credits(enabled = TRUE, # add credits
             text = "Washington Post",
             href = "https://github.com/washingtonpost/data-police-shootings",
             align = "left") %>% 
  hc_add_theme(hc_theme_gridlight())%>%
  hc_legend(enabled = FALSE))
```

### Scatterplot
```{r, echo = T}

(scatter <- highchart() %>%
  hc_add_series_scatter(phl$emp, phl$pax) %>%
  hc_yAxis(title = "PHL Monthy Pax") %>%
  hc_xAxis(title = "PHL Area Monthly Employment")%>%
  hc_title(text = "PHL Employment vs. Monthly Pax", align="left") %>%
  hc_add_theme(hc_theme_gridlight()))

```

### Time Series plot

```{r, echo = T}
library(xts)
pax.ts <- ts(phl$pax, frequency = 12, start = c(2007, 1), end = c(2015,8))

hchart(pax.ts) %>%
  hc_title(text = "Monthly PHL Passengers", align = "left") %>%
  hc_subtitle(text = "January, 2007 to August, 2015", align = "left")%>%
   hc_exporting(enabled = TRUE) %>%
   hc_credits(enabled = TRUE, # add credits
             text = "T-100 Domestic Market (All Carriers)",
             href = "http://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=310&DB_Short_Name=Air%20Carriers") %>%
  hc_add_theme(hc_theme_gridlight())
```

### ACF plot

```{r, echo = T}

acf <- acf(diff(pax.ts), plot = FALSE)

hchart(acf) %>%
  hc_title(text = "ACF Plot of Monthly PHL Passengers", align = "left") %>%
  hc_subtitle(text = "January, 2007 to August, 2015", align = "left") %>%
  hc_add_theme(hc_theme_gridlight()) %>%
  hc_legend(enabled = FALSE)  %>%
   hc_yAxis(title=" ")

```

### Multivariate TS Plot

```{r, echo = T}
pax.ts <- ts(phl$pax, frequency = 12, start = c(2007, 1), end = c(2015,8))
price.ts <- ts(phl$price, frequency = 12, start = c(2007, 1), end = c(2015,8))
earnings.ts <- ts(phl$earnings, frequency = 12, start = c(2007, 1), end = c(2015,8))
emp.ts <- ts(phl$emp, frequency = 12, start = c(2007, 1), end = c(2015,8))

mv.ts<- cbind(pax.ts/1000000, price.ts, earnings.ts/1000, emp.ts/1000000)

hchart(mv.ts) %>%
  hc_title(text = "PHL Passengers, Employment, Earnings, Nat'l Fuel Prices", align = "left") %>%
  hc_subtitle(text = "Monthly Data, January, 2007 to August, 2015", align = "left") %>% 
  hc_add_theme(hc_theme_gridlight()) %>%
  hc_legend(enabled = T)
```







## dygraph (Time Series Plots Only) {.tabset}

### Multivariate TS Plot w/ Custom Colors

```{r, echo=T}
pax.ts <- ts(phl$pax, frequency = 12, start = c(2007, 1), end = c(2015,8))
price.ts <- ts(phl$price, frequency = 12, start = c(2007, 1), end = c(2015,8))
earnings.ts <- ts(phl$earnings, frequency = 12, start = c(2007, 1), end = c(2015,8))
emp.ts <- ts(phl$emp, frequency = 12, start = c(2007, 1), end = c(2015,8))

mv.ts<- cbind(pax.ts/1000000, price.ts, earnings.ts/1000, emp.ts/1000000)

dygraph(mv.ts, main = "Multivariate Time Series Plot") %>% dyRangeSelector()  %>%
  dyOptions(colors = RColorBrewer::brewer.pal(4, "Set1"))

```

### Multivariate TS Shaded Plot W/ Event Line
```{r, echo=T}
pax.ts <- ts(phl$pax, frequency = 12, start = c(2007, 1), end = c(2015,8))
price.ts <- ts(phl$price, frequency = 12, start = c(2007, 1), end = c(2015,8))
earnings.ts <- ts(phl$earnings, frequency = 12, start = c(2007, 1), end = c(2015,8))
emp.ts <- ts(phl$emp, frequency = 12, start = c(2007, 1), end = c(2015,8))

mv.ts<- cbind(pax.ts/1000000, price.ts, earnings.ts/1000, emp.ts/1000000)

dygraph(mv.ts[,1:2], main = "Monthly PHL Pax (Millions) and Nat'l Fuel Prices") %>% dyRangeSelector()%>%
  dyOptions(stackedGraph = TRUE) %>%
  dyRangeSelector(height = 20) %>%
  dyEvent("2008-07-01", "Start of Great Recession", labelLoc = "bottom")

```

###  TS Plot w/ Custom Moving Avg
```{r, echo=T}
pax.ts <- ts(phl$pax/1000, frequency = 12, start = c(2007, 1), end = c(2015,8))

dygraph(pax.ts, main="PHL Pax W/ Custom Moving Avg", ylab = "Monthly Passengers (000s)") %>% dyRangeSelector()%>%
  dyRoller(rollPeriod = 1)

```

###  TS Plot w/ Shading, Annotation and "Follow" Legend
```{r, echo=T}
emp.ts <- ts(phl$emp, frequency = 12, start = c(2007, 1), end = c(2015,8))

dygraph(emp.ts, main="PHL Area Monthly Employment", ylab="Monthly Employment") %>% dyRangeSelector()%>% 
  dyShading(from = "2008-07-01", to = "2009-06-01") %>%
  dyAnnotation("2008-12-01", text = "Great Recession", attachAtBottom = TRUE, width = 120) %>%
  dyLegend(show = "follow")

```

### TS Plot w/ Actual & Predicted
```{r, echo=T}
pax.ts <- ts(phl$pax/1000, frequency = 12, start = c(2007, 1), end = c(2015,8))

pax.hw <- HoltWinters(pax.ts)
pred <- predict(pax.hw, n.ahead = 36, prediction.interval = TRUE)
all <- cbind(pax.ts, pred)

dygraph(all, main="PHL Monthly Pax", ylab = "Monthly Passengers (000s)") %>%
  dySeries("pax.ts", label = "Actual") %>%
  dySeries(c("pred.lwr", "pred.fit", "pred.upr"), label = "Predicted")

```



## Plotly {.tabset}

guide for using with R - https://plot.ly/r/  

using with ggplot2 - https://plot.ly/ggplot2/  

cheatsheet - https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf

### Boxplot
```{r, echo=T}
race2 <- data %>% dplyr::filter(year == "2015") %>% group_by(race2) %>% summarise(total = n())

(p <- plot_ly(na.omit(subset(data, data$age > 0)), y=age, x = race2, color = race2, type = "box") %>% 
  layout(title = "Distribution of Victim Age by Race", xaxis = list(title="Race"), yaxis = list(title="Age")) %>% 
  layout(showlegend = FALSE))

```


### Bubble Chart
```{r, echo = T}
data2 <- data %>% group_by(date, race2) %>% summarise(total = n())

plot_ly(data2, x = date, y = total, text = paste("Fatal Shootings ", total),
        mode = "markers", color = race2)

```

